\TheTitle

### 2015-08-17 linvis.tex
Edits: citations:
Re-wording a bunch of things and adding citations to the paper.

% Extra support
\usepackage{xspace}
\usepackage{caption}

@@

\title{\TheTitle}
\author{
\IEEEauthorblockA{\TheAuthors}
\IEEEauthorblockN{Department of Computer Science,
                    University of Victoria, Canada.}
\IEEEauthorblockA{Email: \TheEmails}
}
\maketitle
\begin{abstract}

	With an average of over 900 top-level merges into the Linux kernel per
	release, some containing thousands of commits, many containing hundreds
	of commits, maintenance of older versions of the kernel becomes nearly
	impossible. For security, performance, and changing hardware needs,
	maintainers must understand the changes made to current versions of the
	kernel, and understand how these changes fit into older version in
        order to make the necessary merges and modifications to the older
        versions. Current tools provide information about repositories through
        the directed acyclic graph (DAG) which is helpful for smaller projects,
        but with the scale and number of branches in the kernel, the DAG
        becomes overwhelming very quickly.

	In this paper, we present a tool that uses the dataset collected by
	German et al. to build a tool that uses a vine model instead of the
	DAG. This tool is designed to cater to the needs of users looking for a
	top-down approach, or a bottom-up approach to navigating the commit
	information of the kernel.
\end{abstract}

@@

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are of the scale of the kernel. Tools like Gitk provide users with the
full view of the DAG, showing all merges and commits. For smaller projects,
this information is useful for showing the relation between various branches.
With large modular projects, like the Linux Kernel, the DAG becomes a tangled
mess (\ref{fig:gitk}) of commits, merges, and the links between them, making it
difficult to derive an explanation of the changes.  Github provides a DAG view
for most projects, but is unable to display the DAG of projects as large as the
Linux Kernel.

@@

Git log, Gitk, and other similar tools only store the authored date of a
commit. A commit may be merged into the project at any time, regardless of when
the commit was authored. Because of this, performing date-range queries does
not provide meaningful results. Performing a date-range query using other tools
over the span of a release will contain results that are for previous versions
of the kernel, and may not contain all of the results for that kernel if
patches were merged after the release date.

German et al. recorded the metadata of every commit and merge in the Linux git
repository from 2012 to 2015, including data that is not stored by git. We are
able to use this information to generate sets of trees instead of a DAG. The
root of each tree is the top-level merge performed by Linus, which merges a
given commit into the kernel.

In this paper, we present the design decisions behind a web-based tool built
around a tree-based model instead of the DAG.

We demonstrate the advantages of this model over the DAG two main use-cases.
Our tool provides information about the location of a given commit in the
respective merge tree, the files edited, the modules edited, and the commit
message in a given commit or merge. The tool allows users to apply various
filters, including the release, a keyword or phrase from the log preview, the
name of the author, or the commit id. The user can request all the top-level
merges containing a commit or merge that matches the query, or all commits and
merges that match the query.

Our core model provides many advantages to other tools. Breaking the data by
merge tree removes the commits and merges of other components of the kernel
from the visualization, providing a clearer picture of what is happening in a
given merge. This model provides additional support to edited file information.

All top-level merges to a given version of the kernel are merged between the
start and release date. Our tree construction allows us to use the top-level
merges to find all merges and commits that have an effect on a given version of
the kernel. The model allows us to better provide mechanisms for performing
date-range queries, providing access to all commits that have an effect on a
given release, even if the commit comes before the start date, or the release
date.

@@

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap design (\ref{fig:treemap}) for
displaying all the file structure of a given version. Treemaps are good for
displaying the topology of shallow, wide trees, making it a good candidate for
displaying file systems. Treemaps are limited in what information can be
displayed. The information within the treemap does not provide clear insight on
what changes were made to the files in the given release. The treemap
visualization only provides information on where the file is located within the
filesystem. File systems and the merge tree of the kernel share many
similarities, in both cases, the representative trees are wide and shallow,
suggesting that a treemap could potentially work to visualize the commit and
merge structure of the kernel.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
	\caption{Treemap of Linux 2.5.33 file structure}
	\label{fig:treemap}
\end{figure}

\subsection{List viewer}
Moghaddam designed a web-based visualization tool displaying the chain of
merges into the kernel. There is no information on what commits are within a
given merge, or what files were edited. The visualization tool provides
information about who authored the commit and when it was authored.
\evantodo{Link to tool}
%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
%% Find paper for this maybe?

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/gitvis.png}
	\caption{List Viewer}
	\label{fig:listviewer}
\end{figure}

\section{Core Model}
Each merge may contain zero or more children of other merges or commits.
Commits have no children, and therefore are always leaf nodes, but commits
contain information about what files are edited, how many lines are edited, and
other important information.

@@

@@

\evantodo{Put this in discussion section}
The main issue with this design is that a user will always move from the
top-down without any additional filtering. Furthermore, it presents the user
with the commit hashes, which provide no explanation of what the commit or
merge may contain.

@@

There are multiple reasons for building the system into a web-based tool.
Putting the tool onto the web enables users to use the system without having to
install anything, making the system platform independent. A web-based tool can
utilize more interactive means of navigation, and provide a better explanation
of what each commit or merge may contain.

The web-based tool provides many mechanisms to enable users to better navigate
and understand changes in the kernel. These mechanisms include searching,
commit messages, files edited, modules edited, and two types of tree viewers.

@@

\begin{verbatim}
current = cid
breadcrumbs = [cid]
while current is not top-level merge:
        breadcrumbs.append(current.next)
return reverse(breacrumbs)
\end{verbatim}

@@

\section{Future Work}

This project was implemented over the course of 4 months, interleaved with
other course projects and assignments. There are still many areas that could
use improvements before the full potential of our model can be reached and
understood with our tool. Here, we outline various areas that still need more
attention.

@@

\evantodo{User cases integrated into files}

% There is currently no functionality surrounding searching for a given file.
% Users who know what file they are working with
%
% We have limited search functionality regarding actual files involved in
% commits. If a user knows which file they are working with and what has changed
% in a given file, they are unable to look specifically for commits pertaining to
% that file. Adding a search filter for searching by file would allow maintainers
% to perform this search.

@@

Currently, there are still various features that have bad performance. More
work is required to get those features responding more quickly, or providing an
asynchronous mechanism and loading bars to improve user experience.

\end{document}

>>>>

% Extra support
\usepackage{xspace}
\usepackage{caption}
\usepackage[nospace]{cite}

@@

\title{\TheTitle}
\author{
\IEEEauthorblockA{\TheAuthors}
\IEEEauthorblockN{Department of Computer Science,
                    University of Victoria, Canada.}
\IEEEauthorblockA{Email: \TheEmails}
}
\maketitle
\begin{abstract}
	With an average of over 900 top-level merges into the Linux kernel per
	release, some containing thousands of commits, many containing hundreds
	of commits, maintenance of older versions of the kernel becomes nearly
	impossible. For security, performance, and changing hardware needs,
	maintainers must understand the changes made to current versions of the
	kernel, and understand how these changes fit into older version in
        order to make the necessary merges and modifications to the older
        versions. Current tools provide information about repositories through
        the directed acyclic graph (DAG) which is helpful for smaller projects,
        but with the scale and number of branches in the kernel, the DAG
        becomes overwhelming very quickly.

	In this paper, we present a tool that uses the dataset collected by
	German et al. to build a tool that uses a tree-based model instead of
        the DAG. This tool is designed to cater to the needs of users looking
        for a top-down approach, or a bottom-up approach to navigating the
        commit information of the kernel.
\end{abstract}

@@

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are of the scale of the kernel. Tools like Gitk provide users with the
full view of the DAG, showing all merges and commits. For smaller projects,
this information is useful for showing the relation between various branches.
The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges, and the
links between them with large modular projects, like the Linux kernel. This
makes it difficult to derive an explanation of the changes made. In other
cases, the kernel is simply too large for the system to generate a
visualization for. Github provides a DAG view for most projects, but is unable
to display the visualization for projects as large as the Linux kernel.

@@

Git log, Gitk, and other similar tools only store the authored date of a
commit. A commit may be merged into the project at any time, regardless of when
the commit was authored, making date-range queries meaningless. Performing a
date-range query using other tools over the span of a release will contain
results that are for previous versions of the kernel, and may not contain all
of the results for that kernel if patches were merged after the release date.

German et al. recorded the metadata of every commit and merge in the Linux git
repository starting in 2011 \cite{German}, including chronological data that is
not stored by git. We are able to use this information to generate sets of
trees instead of a DAG. The root of each tree is the top-level merge performed
by Linus, which merges branches or sets of commits into the kernel.

In this paper, we present the design decisions behind a web-based
tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
model instead of the DAG.  We demonstrate the advantages of this model over the
DAG two main use-cases.  Our tool provides information about the location of a
given commit in the respective merge tree, the files edited, the modules
edited, and the commit message in a given commit or merge.  The tool allows
users to apply various filters, including the release, a keyword or phrase from
the log preview, the name of the author, or the commit id. The user can request
all the top-level merges containing a commit or merge that matches the query,
or all commits and merges that match the query.

Our core model provides advantages to the DAG model used by other tools.
Breaking the commits and merges by merge tree removes the commits and merges of
other components of the kernel from the visualization, providing a clearer
picture of what is happening in a given merge. This model provides additional
support to edited file information. Furthermore, we are able to understand the
changes being made in the order they are brought into the kernel.

All top-level merges to a given version of the kernel are merged between the
start and release date. Our tree construction allows us to use the top-level
merges to find all merges and commits that have an effect on a given version of
the kernel. The model allows us to better provide mechanisms for performing
date-range queries, providing access to all commits that have an effect on a
given release, even if the commit comes before the start date, or after the
release date.

@@

To our knowledge, no repository visualization systems using a tree-based model
are being used. This may be due to the fact that more information than what
is stored by the normal tools is required to create the trees, so servers must
sit and watch as the repository changes over time. Many of the visualization
tools are designed for a general-purpose approach, where any repository is
directly usable, without additional information. Polling for this additional
information makes the application specific to a single repository, or project.

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap visualization
tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
pictured in figure \ref{fig:treemap}, for
displaying all the file structure of a given version. Treemaps are good for
displaying the topology of shallow, wide trees, making it a good candidate for
displaying file systems. Treemaps are limited in what information can be
displayed. The information within the treemap does not provide clear insight on
what changes were made to the files in the given release. The treemap
visualization only provides information on where the file is located within the
filesystem. File systems and the merge tree of the kernel share many
similarities, in both cases, the representative trees are wide and shallow,
suggesting that a treemap could potentially work to visualize the commit and
merge structure of the kernel.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
	\caption{Treemap of Linux 2.5.33 file structure}
	\label{fig:treemap}
\end{figure}

\subsection{List viewer}
Moghaddam designed a web-based visualization
tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
displaying the chain of merges into the kernel. There is no information on what
commits are within a given merge, or what files were edited. The visualization
tool provides information about who authored the commit and when it was
authored. We can see the visualization tool in figure \ref{fig:listviewer}
displaying a part of the kernel commit information, from commit ``26b23ac''
according to the tool.

%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
%% Find paper for this maybe?

\begin{figure}[h]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/gitvis.png}
	\caption{List Viewer}
	\label{fig:listviewer}
\end{figure}

@@

\subsection{Core Model}
Each top-level merge into the kernel is the root-node to a tree. Each merge
may contain zero or more children of other merges or commits. Commits have no
children, and therefore are always leaf nodes, but commits contain information
about what files are edited and how many lines are edited, along with what
modules are involved in a merge.

@@

There are two issues with this design. One issue with this design is that a
user will always move from the top-down without any additional filtering. This
only serves one of our use-cases. The second issue being that it presents the
user with the commit hashes, providing no explanation of what the commit or
merge may contain.

@@

There are multiple reasons for building the system into a web-based tool.
Putting the tool on the web enables users to use the system without having to
install any software, making the system more accessible and platform
independent.A web-based tool can utilize more interactive means of navigation,
and provide a better explanation of what each commit or merge may contain than
the command-line variant.

Our implementation of the web-based tool provides many mechanisms to enable
users to better navigate and understand changes in the kernel. These mechanisms
include searching, commit messages, files edited, modules edited, and two types
of tree viewers.

@@

\begin{verbatim}
current = cid
breadcrumbs = [cid]
while current is not top-level merge:
        breadcrumbs.append(current.next)
	current = current.next
return reverse(breacrumbs)
\end{verbatim}

@@

\section{Future Work}
This project was implemented over the course of 4 months, interleaved with
other course projects and assignments. There are still many areas that could
use improvements before the full potential of our model can be reached and
understood with our tool. Here, we outline various areas that still need more
attention.

@@

There is currently no functionality surrounding searching by filename. Use-case
2 may know what file they are editing and try to determine how it may work with
the other commits and merges in the module. It is possible that a third
use-case may arise, where the user wants to determine all commits that effect a
given file. In both cases, a bottom-to-top approach is used. With this
functionality, implementing the top-to-bottom search feature becomes trivial.

There is limited functionality to the presentation of the file information. At
a minimum, the patches for the commits can be displayed. The patches can then
be used for many more additional features. The patches can be used to piece
together parts of the file to generate a full patch at a given merge rather
than displaying each patch individually. The patches can also be used for
determining what kinds of changes were made, if the lines are being added,
removed, or replaced.

@@

\subsection{Performance}
The modules and list tree view still use the old method of gathering the
information for the respective view type. For single commits or very small
merges, getting everything directly from the database and performing everything
server-side performs better than using the asynchronous requests, as there
isn't the overhead of multiple round trips; however, as the size of the merges
grow, the performance becomes far worse than the asynchronous system and in
some cases will time out. The asynchronous approach adds some overhead,
profiled at roughly 500ms, regardless of the response size. The asynchronous
approach has many benefits, over building the response on the server, including
the ability to add spinners, which is an improvement to user-experience over
looking at a blank page. Further work can be done to improve the results beyond
the response time of 500ms using improved queries and more efficient methods of
building the tree.

\section{Conclusion}

\evantodo{DO IT! JUST DO IT!}

\bibliographystyle{plain}
\bibliography{citations}

% \balance
\end{document}
